---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---
```{r setup}
# set global chunk options: 
library(knitr)
opts_chunk$set(cache=FALSE, fig.align='center')
```
```{r alllibraries, echo=FALSE}
library(ggplot2)
```

## Loading and preprocessing the data
Unzip the file to data folder.
```{r}

data <- "./activity.zip"
csv <- "./data/activity.csv"
unzip(data,exdir="./data")
activity <- read.csv(csv)
activity$date <- as.Date(activity$date,"%Y-%m-%d")
activity.complete_cases <- activity[complete.cases(activity),]
```
## What is mean total number of steps taken per day?

```{r}

stepsPerDay <- aggregate(steps ~ date, data=activity.complete_cases,sum)
hist(stepsPerDay$steps,xlab="Total number of steps per day",main="Histogram of Total number of steps taken each day") 



```

Mean and Median of the total number of steps taken per day

```{r echo=FALSE}
c("Mean" = mean(stepsPerDay$steps),"Median" = median(stepsPerDay$steps))
```

## What is the average daily activity pattern?
```{r}
stepsByInterval <- aggregate(steps ~ interval, data=activity.complete_cases,mean)
plot(stepsByInterval,type="l",xlab="Interval", ylab="Average Steps of All Days", main
="Average Steps of All days by interval")

```


```{r, echo=FALSE}
active <- stepsByInterval[which.max(stepsByInterval[,2]),]
#Formating 
hhmm <- formatC(active[,"interval"],width=4,format="d",flag="0")
most_active_interval <- format(strptime(hhmm,format="%H%M"),"%H:%M")
```
 
    On average across all the days in the dataset, interval `r most_active_interval` contains the maximum number of steps.

## Imputing missing values
 Total number of missing values in the dataset are `r sum(is.na(activity$steps)) `
 
 Imputing Strategy:
 
 Imputing missing values in an interval using the average steps of all the days in that interval.
 
```{r}
activity.imputed_nas <- activity
intervals <- unique(activity$interval)
for (interval in intervals){
  activity.imputed_nas[which(activity$interval==interval & is.na(activity$steps)),"steps"] <-
    stepsByInterval[which(stepsByInterval$interval==interval),"steps"]
}
stepsPerDay.imputes_nas <- aggregate(steps ~ date, data=activity.imputed_nas,sum)
```

```{r}
hist(stepsPerDay.imputes_nas$steps,xlab="Total number of steps per day",main="Histogram of Total number of steps taken each day with imputed missed values") 

```
After imputing the missing values, the Mean and the Median of the total number of steps taken per day

```{r,echo=FALSE}
c("Mean" = mean(stepsPerDay.imputes_nas$steps),"Median" = median(stepsPerDay.imputes_nas$steps))
```
> There is a negligible change in the median, and mean after imputing the missing values.

## Are there differences in activity patterns between weekdays and weekends?

```{r fig.width=7, fig.height=6}


activity.imputed_nas$wday <- as.POSIXlt(activity.imputed_nas$date)$wday

activity.imputed_nas$weekdayorweekend <- "weekday"

# 0=sunday,6 = saturday
activity.imputed_nas[which(activity.imputed_nas$wday ==6 | activity.imputed_nas$wday ==0),]$weekdayorweekend <- "weekend"


activity.imputed_nas <-transform(activity.imputed_nas,weekdayorweekend=factor(weekdayorweekend) )
avergastepsPerInterval.weekendorweekday <- aggregate(steps ~ interval + weekdayorweekend, data=activity.imputed_nas,mean)

qplot(interval, steps ,data=avergastepsPerInterval.weekendorweekday , 
       geom=c("point","line") ,
      facets= . ~ weekdayorweekend,
     main="Average steps of all days in an interval categorized weekday or weekends",
     xlab="Interval",
     ylab="Average steps of all days")
```